---
title: "Data Merging and Summary Statistics"
author: "Abay, Anthony, Daniel, Quinton"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ACLED Data

We will begin by reading in the data from ACLED on armed conflicts during the Arab spring. This data will be cleaned and aggregated since the factors we are looking at are at the country level.

```{r}
all_countries = read.csv('all_countries.csv', stringsAsFactors =TRUE)

#selecting just the necessary countries for the project since the website is too slow to query
arab_spring = c('Libya', 'Egypt', 'Yemen', 'Syria', 'Bahrain', 'Morocco', 'Iraq','Algeria',
                'Jordan', 'Kuwait', 'Oman', 'Sudan', 'Lebanon', 'Tunisia')

protest_type = c('Riots', 'Protests', 'Violence against civilians', 'Explosions/Remote Violence')#reducing to the relevant types of protest
all_protests = subset(all_countries, event_type %in% protest_type)

countries = subset(all_protests, country %in% arab_spring)#creating the data.frame of the right countries
countries$country = droplevels(countries$country)

countries = subset(countries, select = c('event_date', 'year','event_type','country','location',
                                         'latitude','longitude'))#non aggregated data with correct columns

#write.csv(countries, 'countries.csv')

#aggregating the data by country
agg_countries = data.frame()
relevant_years = c(2005, 2006, 2007, 2008, 2009, 2010,2011,2012)

for (temp_country in unique(countries$country)){
  for (years in relevant_years){
    temp = subset(countries, country == temp_country & year== years)
    total = nrow(temp)
    riot_count = sum(temp$event_type=='Riots')
    explosion_count = sum(temp$event_type=='Explosions/Remote violence')
    protest_count = sum(temp$event_type=='Protests')
    vac_count = sum(temp$event_type=='Violence against civilians')
    temp_list = c(temp_country, years, total, riot_count, explosion_count, protest_count, vac_count)
    agg_countries = rbind(agg_countries, temp_list) 
  }
}

names(agg_countries) = c('country', 'year','total_events','riot_count', 'explosion_count', 'protest_count',
                         'vac_count') #vac_count stands for violence against citizens

#ordering the data to make binding easier
pacman::p_load(plyr)
agg_countries = arrange(agg_countries, country, year)

#none of the countries we are looking at had any explosions
agg_countries = subset(agg_countries, select = -c(explosion_count))

#we also want our counts as integers
agg_countries = transform(agg_countries, year = as.integer(year),total_events = as.integer(total_events), riot_count = as.numeric(riot_count), protest_count = as.integer(protest_count), 
                          vac_count = as.integer(vac_count))#converting to appropriate data types
```

Next we will merge the datasets that contain the information on our factors.

# CPI data

```{r}
arab_spring = unique(agg_countries$country)#resetting this vector since we only have data on 6 countries

cpi_data = read.csv('CPI.csv', stringsAsFactors = TRUE)
cpi_data = subset(cpi_data, select = c(country, score, rank, year))#the region won't be relevant to our analysis

#subsetting the correct countries
cpi_data = subset(cpi_data, country %in% arab_spring)
cpi_data = arrange(cpi_data, country,year)
cpi_data = subset(cpi_data, select = c(score, rank))#getting the new data to add

#converting score to a numeric
cpi_data = transform(cpi_data, score = as.numeric(score))

agg_countries = cbind(agg_countries,cpi_data)#binding it to the existing dataframe
```

# Education Data
The data that ranks a countries education levels is in a time series so we will need to do more wrangling with it.

```{r}
education = read.csv('Education.csv')
education = subset(education, Country %in% arab_spring)#getting the relevant countries
education = subset(education, select = c(HDI.Rank..2018., Country,X2005, X2006, X2007, X2008, X2009, X2010, X2011, X2012))#getting the relevant years

#converting to numeric
education = transform(education, X2005 = as.numeric(X2005),X2006 = as.numeric(X2006),X2007 = as.numeric(X2007), X2008 = as.numeric(X2008),X2009 = as.numeric(X2009), X2010 = as.numeric(X2010), X2011 = as.numeric(X2011), X2012 = as.numeric(X2012))

names(education) = c("HDI_rank", "country","2005","2006","2007","2008","2009", "2010",'2011','2012')#renaming columns
education = arrange(education, country)

HDI_score = c()#this list will become the columns containting the hdi scores on education
for (i in 1:nrow(education)){#need to run for every row
  for (j in 3:10){#need to run for the last three columns which contain the data from the years
    HDI_score = append(HDI_score, education[i,j])
  }
}
agg_countries = cbind(agg_countries, HDI_score)
```

# GDP per Capita

```{r}
gdp = read.csv('GDP_per_capita.csv')
arab_spring = append(arab_spring, 'Egypt, Arab Rep.')#Egypt is listed as "Egypt, arab rep." in the worldbank
gdp = subset(gdp, Country.Name %in% c(arab_spring))#getting the relevant countries
gdp = subset(gdp, select = c(Country.Name, X2005, X2006, X2007, X2008, X2009, X2010, X2011, X2012))#getting the relevant years

GDP_per_capita = c()#this list will become the columns containting the hdi scores on education
for (i in 1:nrow(gdp)){#need to run for every row
  for (j in 2:9){#need to run for the last three columns which contain the data from the years
    GDP_per_capita = append(GDP_per_capita, gdp[i,j])
  }
}
agg_countries = cbind(agg_countries, GDP_per_capita)
```

# Unemployment

```{r}
unemployment = read.csv('unemployment.csv')
unemployment = subset(unemployment, Country.Name %in% c(arab_spring))#getting the relevant countries
unemployment = subset(unemployment, select = c(Country.Name, X2005, X2006, X2007, X2008, X2009, X2010, X2011, X2012))#getting the relevant years

unemployment_percent = c()
for (i in 1:nrow(gdp)){#need to run for every row
  for (j in 2:9){#need to run for the last three columns which contain the data from the years
    unemployment_percent = append(unemployment_percent, unemployment[i,j])
  }
}
agg_countries = cbind(agg_countries, unemployment_percent)
```

# Final data set

```{r}
write.csv(agg_countries, file='final_data.csv')#writing the data set to be saved
head(agg_countries)#there aren't many rows for observations so we can view the whole thing
summary(agg_countries)

#using histograms to get a brief insight on distributions
for (i in 3:ncol(agg_countries)){
  hist(agg_countries[,i], main = colnames(agg_countries[i]),
       xlab = colnames(agg_countries[i]))
}
```
